Unbalanced Mallows Models for Optimizing Expensive Black-Box Permutation Problems

### 2020-10-30 paper/main.tex
minor

The proposed UMM drastically differs from EDAs since EDAs assume that function
evaluations are cheap and they can do many of them. On the contrary, UMM aims
at learning as much as possible with a limited number of function
evaluations. Moreover, unlike EDAs, UMM does not rely on a fixed-size
population of solutions that evolves over ``generations''. Instead, taking
inspiration from Bayesian optimizers, UMM considers a sample of permutations
during the whole execution that is increased by one new permutation at each
iteration. Finally, the balance between intensification and exploration is
dynamically adapted by controlling the variance of the probabilistic Mallows
model.

@@

\manuel{}{
  \begin{itemize}
  \item Who else has used Mallows model for black-box optimization? Ekhine: Va a mirarlo.
  \item Who else has used unbalanced borda for optimization? Ekhine va a añadir 2 frases.
  \end{itemize}
}
This paper structured as follows. Section~\ref{sec:backgroud} summarizes fundamental concepts about permutation spaces, the uBorda algorithm, black-box combinatorial optimization under a limited budget and CEGO, a Bayesian optimizer for such problems. Section~\ref{sec:umm} presents the UMM algorithm, which is the main contribution of this paper. UMMA is analyzed using the experimental setup presented in Section~\ref{sec:setup} and results are presented in Section~\ref{sec:analysis}. We end the paper with overall conclusions and a call for further research on this type of optimization scenario (Section~\ref{sec:conclusions}).

@@

A somewhat surprising result was that the behavior of UMM (and the same for
CEGO) on the PFSP is close to random. This result may be due to the distance
metric used, but further analysis is needed to corroborate this
conjecture. Here, we have focused on Kendall's $\tau$ distance, however, it is
possible to extend UMM to other distance metrics, which will allow us to
dynamically select among various distance metrics for an unknown black-box \ekhine{}{estoy totalmente en contra de esto :) }
permutation landscape~\citep{ZaeStoBar2014:ppsn}. Also, as shown by our
experiments, the behavior of the parameters ($r_1$ and $r_2$) of UMM is not
always obvious, thus a more detailed analysis would be needed to provide either
generally good static values or an online adaptation approach. Finally, we have
used here three very different combinatorial problems (LOP, PFSP and QAP) as
black-box benchmarks. However, an even more diverse range of problems would be
needed to understand the behavior of UMM on real-world black-box combinatorial
landscapes.

>>>>

The proposed UMM drastically differs from EDAs since EDAs assume that function
evaluations are cheap and they can do many of them. On the contrary, UMM aims
at learning as much as possible with a limited number of function
evaluations. Moreover, unlike EDAs, UMM does not rely on a fixed-size
population of solutions that evolves over ``generations''. Instead, taking
inspiration from Bayesian optimizers, UMM considers a sample of permutations
during the whole execution that is increased by one new permutation at each
iteration.
%
Finally, the balance between intensification and exploration is dynamically
adapted by controlling the variance of the probabilistic Mallows model.

@@

\manuel{}{
  \begin{itemize}
  \item Who else has used Mallows model for black-box optimization? Ekhine: Va a mirarlo.
  \item Who else has used unbalanced borda for optimization? Ekhine va a añadir 2 frases.
  \end{itemize}
}

This paper structured as follows. Section~\ref{sec:backgroud} summarizes fundamental concepts about permutation spaces, the uBorda algorithm, black-box combinatorial optimization under a limited budget and CEGO, a Bayesian optimizer for such problems. Section~\ref{sec:umm} presents the UMM algorithm, which is the main contribution of this paper. UMM is analyzed using the experimental setup presented in Section~\ref{sec:setup} and results are presented in Section~\ref{sec:analysis}. We end the paper with overall conclusions and a call for further research on this type of optimization scenario (Section~\ref{sec:conclusions}).

@@

A somewhat surprising result was that the behavior of UMM (and the same for
CEGO) on the PFSP is close to random. This result may be due to the distance
metric used, but further analysis is needed to corroborate this
conjecture. Here, we have focused on Kendall's $\tau$ distance, however, it is
possible to extend UMM to other distance metrics, which will allow us to
dynamically select among various distance metrics for an unknown black-box
% \ekhine{}{estoy totalmente en contra de esto :) }
permutation landscape~\citep{ZaeStoBar2014:ppsn}. Also, as shown by our
experiments, the behavior of the parameters ($r_1$ and $r_2$) of UMM is not
always obvious, thus a more detailed analysis would be needed to provide either
generally good static values or an online adaptation approach. Finally, we have
used here three very different combinatorial problems (LOP, PFSP and QAP) as
black-box benchmarks. However, an even more diverse range of problems would be
needed to understand the behavior of UMM on real-world black-box combinatorial
landscapes.
